A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-06-09 main.tex
Improve Introduction.

Multistage programming can also optimize vector calculation.
For example, \verb|add| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|add| function for a given vector length and optimize it with multistage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

@@

However, there is a sever problem in functions which are optimized by multistage programming.
Although unrolled \verb|add| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.

@@

Dependent types are types which is dependent on values.
We can use dependent types for securer programming than ordinary types.
For example, we can realize vectors with their sizes with them.
\begin{verbatim}
        Vector :: Int -> *
\end{verbatim}
\verb|Vector| is a type constor which takes the length of vectors.
\verb|Vector 3| is a type for vectors whose lengths are 3.
We can confirm the arguments of \verb|add| function is the same with dependent types.
\red{secure programming? finer grained typing? とか}

\subsection{Multistage Programming with Dependent Types}
\red{このsectionは短いので上とくっつけてもいいかもしれない}

@@

As above, functions generated with multistage programming can take more restricted values as arguments.
When you optimize \verb|add| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multistage calculus
so that the type system can guarantee optimized functions are used properly.

>>>>

Multistage programming can also optimize vector calculation.
For example, \verb|vadd| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|vadd| function for a given vector length and optimize it with multistage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

@@

However, there is a sever problem in functions which are optimized by multistage programming.
Although unrolled \verb|vadd| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.

@@

Dependent types are types which is dependent on values.
We can use dependent types for securer programming.
For example, we can realize vectors with their sizes with them.
\begin{verbatim}
        Vector :: Int -> *
\end{verbatim}
\verb|Vector| is a type constor which takes the length of vectors, that is, 
\verb|Vector 3| is a type for vectors whose lengths are 3.
If \verb|vadd| function has the type of \verb|Vector n -> Vector n -> Vector n|,
we can confirm the two arguments of \verb|vadd| function has the same length.
\red{secure programming? finer grained typing? とか}

\subsection{Multistage Programming with Dependent Types}
\red{このsubsectionは短いので上とくっつけてもいいかもしれない}

@@

As we pointed out in the above section,
functions optimized with multistage programming can take only restricted values as arguments.
When you optimize \verb|vadd| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multistage calculus
so that the type system can guarantee optimized functions are used properly.

In this paper, we design new multistage calculus \LMD by 
merging dependent types into existing multistage calculus \LTP\cite{Hanada2014}.
\LMD is multistage calculus which contains dependent types.
\LMD makes multistage programming more safer with the power of dependent types.
